---
title: "Gov 50 Final Project"
author: Jada Pierre
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

I am interested in exploring this data to investigate if there is a relationship among the citizen voting age for communities of color. I am specifically interested in the Latino and Asian voting population because thy are two of the biggest immigrant population in America. I am interested in looking at migrant voting behavior in the United States. I also find that voting behavior is more visible for other racial and ethnic groups such as Black and white people in comparison to Latinos and Asians. If more people study this data, one can learn to build solidarity between these communities.

## Data and Research Design

I am exploring data from the American Community Survey from the U.S. Census in 2017 revolving around the citizen voting age population and race and ethnicity. The data set includes information regarding U.S. states, more specifically racial demographics and within it, who is eligible to vote. My research question is as follows: What is the relationship between the Latinos and Asians who are able to vote?

My final project will be a observational study. To investigate the relationship, my explanatory variable is the proportion of Latinos that can vote. My independent variable is the non-citizen voting population for Latinos and the non-citizen voting population Asians. My dependent variable is the adult population for Latinos and the adult population for Asians.

For my final project, I made the proportions based off of total estimates. I did this to help examine changes in the variables more easily.

The CVAP proportion of Latinos and Asians is the proportion of those groups that are United States citizens 18 years of age or older that can vote. The ADU_EST represents the adult estimate of Asians and Latinos the total number of people 18 years of age or older in a group. The NON_CIT variable is the proportion of Latinos and Asians that are non-citizens in the United States.


```{r include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center")
library(tidyverse)
library(patchwork)
state <- read.csv("state.csv") |>
  select(!ends_with("MOE"),
         -GEOID, -LNNUMBER) |>
  filter(LNTITLE %in% c("Hispanic or Latino", "Asian Alone")) |>
  mutate(LNTITLE = case_when(LNTITLE == "Hispanic or Latino" ~ "Latino",
                             LNTITLE == "Asian Alone" ~ "Asian"))
```

```{r}
cvap_data <- state |>
  pivot_wider(
    names_from = LNTITLE,
    values_from = c(TOT_EST, ADU_EST, CIT_EST, CVAP_EST)
  ) |>
  mutate(CVAP_prop_Latino = CVAP_EST_Latino/TOT_EST_Latino,
         CVAP_prop_Asian = CVAP_EST_Asian/TOT_EST_Asian,
         ADU_prop_Latino = ADU_EST_Latino/TOT_EST_Latino,
         ADU_prop_Asian = ADU_EST_Asian/TOT_EST_Asian,
         NONCIT_prop_Latino = (TOT_EST_Latino - CIT_EST_Latino)/TOT_EST_Latino,
         NONCIT_prop_Asian = (TOT_EST_Asian - CIT_EST_Asian)/TOT_EST_Asian)
```

### Distribution Plot

```{r fig.width=8, fig.height=4}
latinodistro <- cvap_data |>
  ggplot(aes(x = CVAP_prop_Latino)) +
  geom_histogram(fill = "red") +
  labs(x = "Proportion of Latinos that Can Vote",
       y = "Count",
       title = "Distribution Plot of Proportion of Latinos that Can Vote")

asiandistro <- cvap_data |>
  ggplot(aes(x = CVAP_prop_Asian)) +
  geom_histogram(fill = "blue") +
  labs(x = "Proportion of Asians that Can Vote",
       y = "Count",
       title = "Distribution Plot of Proportion of Asians that Can Vote")

latinodistro + asiandistro
```

### Interpretation

#### Median for CVAP Latino
```{r}
median(cvap_data$CVAP_prop_Latino)
```

#### Median for CVAP Asian
```{r}
median(cvap_data$CVAP_prop_Asian)
```


For the distribution plot of Latinos that can vote, the proportion ~52% appears six times in the dataset. The curve of the plot expands from ~30% to ~63%. There is an outlier at ~78%. Thus, we can expect per state, there will be between ~30-63% voting Latinos. The median expected proportion for any given state can be expected to be 44.2%.

For the distribution plot of Asians that can vote, the proportion ~50% appears ten times in the dataset. The curve of the plot expands from ~30% to ~64%. There is an outlier at ~73%. Thus, we can expect per state, there will be around ~30-64% voting Asians. The median expected proportion for any given state can be expected to be 49.5%.

## Results

### Plot Summary

```{r fig.width=8, fig.height=4}
bivariatelatino <- cvap_data |>
  ggplot(aes(x = NONCIT_prop_Latino,
                     y = CVAP_prop_Latino)) +
  geom_point() +
  geom_smooth(color = "red", se=FALSE, method = "lm") +
  labs(x = "Proportion of Non-Citzens",
       y = "Proportion of Latinos that Can Vote",
       title = "Relationship Between Latino CVAP and Non-Citizen Populations")

bivariateasian <- cvap_data |>
  ggplot(aes(x = NONCIT_prop_Asian,
                     y = CVAP_prop_Asian)) +
  geom_point() +
  geom_smooth(color = "blue", se=FALSE, method = "lm") +
  labs(x = "Proportion of Non-Citzens",
       y = "Proportion of Asians that Can Vote",
       title = "Relationship Between Asian CVAP and Non-Citizen Populations")

bivariatelatino + bivariateasian
```

### Interpretation

Both bivariate plot summaries display linear inverse relationships. The data on both plots is close to the line with very few outliers and variates. As the proportion of non-citizens increases for both Latinos and Asians, the proportion of their voting age population is associated with a decrease.

### Multiple Regression Results

```{r}
latinomultiregresults <- summary(lm(data = cvap_data, CVAP_prop_Latino ~ NONCIT_prop_Latino + ADU_prop_Latino))$coefficients
latinomultiregresults |>
  knitr::kable(digits = 2)
```

```{r}
asianmultiregresults <- summary(lm(data = cvap_data, CVAP_prop_Asian ~ NONCIT_prop_Asian + ADU_prop_Asian))$coefficients
asianmultiregresults |>
  knitr::kable(digits = 2)
```


### Interpretation

For the Latino multiple regression model, the intercept is statistically significant. For every one percent increase in the non-citizen Latino population proportion, there's 0.92% decrease in the Latino citizen voting age population proportion holding all other variables constant. Additionally, for every one percent increase in the adult Latino population proportion, there is a 1% increase in the CVAP holding all other variables constant.

For the Asian multiple regression model, the intercept is not statistically significant. For every one percent increase in the non-citizen Asian population proportion, there is a 0.79% increase in the Asian citizen voting age population proportion holding all other variables constant. Additionally, for every one percent increase in the adult Asian population proportion, there is a 0.15% increase in the CVAP holding all other variables constant.

The Latino non-citizen voting age population proportion slightly lowers the Latino CVAP. In contrast, The Asian non-citizen voting age population proportion increases the Asian CVAP. The Latino adult population proportion increases the Latino CVAP by a 1% increase. In comparison, the Asian adult population only increases the Asian CVAP increases by 0.15% which is slightly lower than the increase in the Latino CVAP.

### Causality and Confounders

A possible confounder that could be affecting the voting age population for both Latinos and Asians in the United States is their education level. I selected this variable because there is a strong relationship between higher education levels and voter turnout. This is affecting the results for the Latino multiple regression because [...]. It also affects the Asian multiple regression because [...].

Another confounder is income status. I selected this variable because low-income individuals are less likely to turnout and vote due to voter suppression. This is affecting the results for the Latino multiple regression because [...]. It also affects the Asian multiple regression because [...].

## Conclusion

### Next Steps

What other researchers could further investigate in this direction is the Latino and Asian voting population on a county level as one can draw more in depth conclusions with this data. While the observations are sorted into racial categories and an ethnicity, it would be beneficial to view the ethnic makeup of these races to prevent from treating these groups as monolith. This would be looking more at specific ethnicities such as Afro-Cubans and White Italians in the United States for example and their voting behavior.