---
title: "Latino and Asian Populations in the United States"
author: Jada Pierre
description: "Researching the relationship between the Latinos and Asians who are able to vote and their populations."
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

I am interested in exploring this data to investigate if there is a relationship among the citizen voting age for communities of color. I am specifically interested in the Latino and Asian voting population because thy are two of the biggest immigrant population in America. I am interested in looking at migrant voting behavior in the United States. I also find that voting behavior is more visible for other racial and ethnic groups such as Black and white people in comparison to Latinos and Asians. If more people study this data, one can learn to build solidarity between these communities.

## Data and Research Design

I am exploring data from the American Community Survey from the U.S. Census in 2017 revolving around the citizen voting age population and race and ethnicity. The data set includes information regarding U.S. states, more specifically racial demographics and within it, who is eligible to vote. My research question is as follows: What is the relationship between the Latinos and Asians who are able to vote and their population?

My final project will be a observational study. To investigate the relationship, my explanatory variable is the proportion of Latinos that can vote. My independent variable is the non-citizen voting population for Latinos and the non-citizen voting population Asians. My dependent variable is the adult population for Latinos and the adult population for Asians.

For my final project, I made the proportions based off of total estimates. I did this to help examine changes in the variables more easily.

The CVAP proportion of Latinos and Asians is the proportion of those groups that are United States citizens 18 years of age or older that can vote. The ADU_EST represents the adult estimate of Asians and Latinos the total number of people 18 years of age or older in a group. The NON_CIT variable is the proportion of Latinos and Asians that are non-citizens in the United States.


```{r include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center")
library(tidyverse)
library(patchwork)
state <- read.csv("state.csv") |>
  select(!ends_with("MOE"),
         -GEOID, -LNNUMBER) |>
  filter(LNTITLE %in% c("Hispanic or Latino", "Asian Alone")) |>
  mutate(LNTITLE = case_when(LNTITLE == "Hispanic or Latino" ~ "Latino",
                             LNTITLE == "Asian Alone" ~ "Asian"))
```

```{r}
cvap_data <- state |>
  pivot_wider(
    names_from = LNTITLE,
    values_from = c(TOT_EST, ADU_EST, CIT_EST, CVAP_EST)
  ) |>
  mutate(CVAP_prop_Latino = CVAP_EST_Latino/TOT_EST_Latino,
         CVAP_prop_Asian = CVAP_EST_Asian/TOT_EST_Asian,
         ADU_prop_Latino = ADU_EST_Latino/TOT_EST_Latino,
         ADU_prop_Asian = ADU_EST_Asian/TOT_EST_Asian,
         NONCIT_prop_Latino = (TOT_EST_Latino - CIT_EST_Latino)/TOT_EST_Latino,
         NONCIT_prop_Asian = (TOT_EST_Asian - CIT_EST_Asian)/TOT_EST_Asian)
```

### Distribution Plots

```{r fig.width=8, fig.height=4}
latinodistro <- cvap_data |>
  ggplot(aes(x = CVAP_prop_Latino)) +
  geom_histogram(aes(y = ..count../sum(..count..)), fill = "skyblue") +
  labs(x = "Latino CVAP Proportion",
       y = "Proportion",
       title = "Distribution of State Level Latino CVAP")

asiandistro <- cvap_data |>
  ggplot(aes(x = CVAP_prop_Asian)) +
  geom_histogram(aes(y = ..count../sum(..count..)), fill = "maroon") +
  labs(x = "Asian CVAP Proportion",
       y = "Proportion",
       title = "Distribution of State Level Asian CVAP")

latinodistro + asiandistro
```

The median expected proportion for any given state can be expected to be 44.2%. The distribution of the outcome ranges from ~30% to ~63%. Thus, we can expect per state, there will be between ~30-63% Latino citizens of voting age.

The median expected proportion for any given state can be expected to be 49.5%. The distribution of the outcome ranges from ~30% to ~64%. Thus, we can expect per state, there will be around ~30-64% Asians of voting age.

While there appears to be no huge difference in the range, there is a slight difference in medians.

## Results

The previous section looked a range of state CVAP estimates and displays how the outcome varies among the main explanatory variable.

### Plot Summary

```{r fig.width=12, fig.height=4}
bivariatelatino <- cvap_data |>
  ggplot(aes(x = NONCIT_prop_Latino,
                     y = CVAP_prop_Latino)) +
  geom_point() +
  geom_smooth(color = "skyblue", se=FALSE, method = "lm") +
  labs(x = "Proportion of Non-Citzens",
       y = "Proportion of Latinos that Can Vote",
       title = "Relationship Between Latino CVAP and Non-Citizen Populations")

bivariateasian <- cvap_data |>
  ggplot(aes(x = NONCIT_prop_Asian,
                     y = CVAP_prop_Asian)) +
  geom_point() +
  geom_smooth(color = "maroon", se=FALSE, method = "lm") +
  labs(x = "Proportion of Non-Citzens",
       y = "Proportion of Asians that Can Vote",
       title = "Relationship Between Asian CVAP and Non-Citizen Populations")

bivariatelatino + bivariateasian
```

### Interpretation

Both bivariate plot summaries display a negative relationship between the proportion non-citizens and their voting age populations. The range for the relationship between Latino CVAP and non-citizen populations is around 0.05-0.35. It is mainly concentrated around 0.15-0.34. The range for the relationship between Asian CVAP and non-citizen populations is around 0.14-0.57. It is mainly concentrated around 0.25-0.45.

The data on both plots is close to the line with very few outliers and variates. As the proportion of non-citizens increases for both Latinos and Asians, the proportion of their voting age population is associated with a decrease.

### Multiple Regression Results

```{r}
latinomultiregresults <- summary(lm(data = cvap_data, CVAP_prop_Latino ~ NONCIT_prop_Latino + ADU_prop_Latino))$coefficients
latinomultiregresults |>
  knitr::kable(digits = 2)
```

```{r}
asianmultiregresults <- summary(lm(data = cvap_data, CVAP_prop_Asian ~ NONCIT_prop_Asian + ADU_prop_Asian))$coefficients
asianmultiregresults |>
  knitr::kable(digits = 2)
```


### Interpretation

For the Latino multiple regression model, the intercept is statistically significant. On average for every one percent increase in the non-citizen Latino population proportion, there's 0.92% decrease in the Latino citizen voting age population proportion holding all other variables constant. Additionally, on average for every one percent increase in the adult Latino population proportion, there is a 1% increase in the CVAP holding all other variables constant.

For the Asian multiple regression model, the intercept is not statistically significant. On average for every one percent increase in the non-citizen Asian population proportion, there is a 0.79% increase in the Asian citizen voting age population proportion holding all other variables constant. Additionally, on average for every one percent increase in the adult Asian population proportion, there is a 0.15% increase in the CVAP holding all other variables constant.

The Latino non-citizen voting age population proportion is associated with a lower Latino CVAP. In contrast, The Asian non-citizen voting age population proportion is associated with a smaller reduction the Asian CVAP. The Latino adult population proportion increases the Latino CVAP by a 1% increase. In comparison, the regression table suggests that, on average, holding all other covariates constant, a one percent increase in the Asian adult population is associated with a 0.79% increase in the Asian CVAP population.

### Causality and Confounders

A possible confounder that could be affecting the voting age population for both Latinos and Asians in the United States is their age. For example, in order to be eligible to vote, one must be a citizen. To be a U.S. citizen through naturalization, one must be at least 18 years or older by the time they submit their application. Moreover, age can possibly affect the non-citizen voting population for Latinos and Asians and the adult population for Latinos and Asians. It could possibly be affecting both variables age is a factor that determines if one is eligible to apply for citizenship or not. Therefore, one cannot control for age as a possible confounder.

## Conclusion

My research displays the relationship between the non-citizen voting population for Latinos and the non-citizen voting population Asians and the adult population for Latinos and the adult population for Asians. Both relationships of interest show a correlation between these two variables. A limitation of this study is that the ethnicities within the Latino and Asian groups remain unknown.

### Next Steps

What other researchers could further investigate in this direction is the Latino and Asian voting population on a county level as one can draw more in depth conclusions with this data. While the observations are sorted into racial categories and an ethnicity, it would be beneficial to view the ethnic makeup of these races to prevent from treating these groups as monolith. This would be looking more at specific ethnicities such as Afro-Cubans and White Italians in the United States for example and their voting behavior.