---
title: "Gov 50 Final Project"
author: Jada Pierre
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

I am interested in exploring data from the American Community Survey from the U.S. Census revolving around voting patterns and race and ethnicity. The dataset includes information regarding state, race and ethnicity, total estimated population (rounded) and its margin of error, and citizenship status for U.S. citizens in that geographic area and group (rounded) and its margin of error.

Thus, my research question is as follows: What is the relationship between the Latinos and Asians who are able to vote?

dependent variable: proportion of citizens
ind. variable: total population
control: proportion over 18 (CVAP)

```{r}
library(tidyverse)
library(dbplyr)
state <- read.csv("state.csv") |>
  select(!ends_with("MOE"),
         -GEOID, -LNNUMBER) |>
  filter(LNTITLE %in% c("Hispanic or Latino", "Asian Alone")) |>
  mutate(LNTITLE = case_when(LNTITLE == "Hispanic or Latino" ~ "Latino",
                             LNTITLE == "Asian Alone" ~ "Asian"))
state
```

```{r}
cvap_data <- state |>
  pivot_wider(
    names_from = LNTITLE,
    values_from = c(TOT_EST, ADU_EST, CIT_EST, CVAP_EST)
  ) |>
  mutate(CVAP_prop_Latino = CVAP_EST_Latino/TOT_EST_Latino) |>
  mutate(CVAP_prop_Asian = CVAP_EST_Asian/TOT_EST_Asian)
cvap_data
```

```{r}
cvap_data |>
  ggplot(aes(x = CVAP_prop_Latino,
                     y = CVAP_prop_Asian)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Total Proportion of Latinos that can vote",
       y = "Total Proportion of Asians that can vote",
       title = "Relationship between Asians and Latinos that can vote")
```

